```---
title: "R ML excercises"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

Load necessary data.

```{r}
check.packages <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
packages<-c("ggplot2", "tidyverse", "lubridate", "caret", "dslabs", "dplyr", "plyr","HistData")
check.packages(packages)

data("heights")
data("iris")
```

Split by predictors and predicted value.

```{r}
y<-heights$sex
x<-heights$height
```

Split tet into training and evaluation.

```{r echo=TRUE}
set.seed(2)
test_index <- createDataPartition(y, times=1, p=0.5, list=FALSE)
test_set <- heights[test_index, ]
training_set <- heights[-test_index, ]

```

Determine y_hat for the test set just by guessing.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
y_hat <- sample(c("Male", "Female"), size = length(test_index), replace=TRUE)
mean(y_hat == test_set$sex)
```
Consider the bias, that males are higher than females.
```{r}
heights %>% group_by(sex) %>% summarize(mean(height), sd(height))
```

Naively predict Male and Female based on height. (;

```{r}
y_hat <- ifelse(x>62, "Male", "Female") %>% factor(levels=levels(test_set$sex))
table(y_hat)
print("Mean accurate answers:")
mean(y==y_hat)
```

Check other vut off values, except for 62:
```{r}
cutoff <- seq(61, 70)
accuracy <- map_dbl(cutoff, function(cutoff_value) {
  y_hat <- ifelse(test_set$height>cutoff_value, "Male", "Female") 
  mean(y_hat == test_set$sex)
})
ggplot(data=data.frame(cutoff=cutoff, accuracy=accuracy), aes(x=cutoff, y=accuracy)) + geom_line() + geom_point() + ggtitle("Accuracy distribution depending on 61:70 cutoff")
```

Check out the accuracy by sex
```{r}
heights %>% mutate(y_hat = y_hat) %>%
  group_by(sex) %>%
  summarize(accuracy = mean(y_hat == sex))
```

Notes on confusion matrixes.
**Sensitivity**: The avility of algo to predict positive outcome, when the actual outcome is positive: P($\widehat{Y}$=1|Y=1). Also Sensitivity = TP/(TP + FN) 
**Specificity**: The avility of algo to **not** predict positive outcome, when the actual outcome is **not** positive: P($\widehat{Y}$=0|Y=0). Also Specificity = TN/(TN + FP) 

```{r}
confusionMatrix(data=y_hat, reference=heights$sex)
```

Get F_1 harmonised score based on the cutoff:
```{r}
F_1 <- map_dbl(cutoff, function(cutoff_value) {
  y_hat <- ifelse(test_set$height>cutoff_value, "Male", "Female") %>% factor(levels=levels(test_set$sex))
  F_meas(data = y_hat, reference = factor(test_set$sex))
})
ggplot(data=data.frame(cutoff=cutoff, F_meas=F_1), aes(x=cutoff, y=F_meas)) + geom_line() + geom_point() + ggtitle("Accuracy distribution depending on 61:70 cutoff")
```
#The test set bias
The test set has more males, so guessing 90% of the time males, would give a mean up to 72-3%:
```{r}
y_hat <- sample(c("Male","Female"), length(test_index), replace=TRUE, prob=c(0.9, 0.1)) %>% factor(levels = levels(test_set$sex))
mean(y_hat==test_set$sex)
```

#Receiver Operating Characteristics - the ROC curve
```{r}
cutoffs <-c(50, seq(60, 75), 80)
height_cutoff <- map_df(cutoffs, function (cutoff_value) {
    y_hat <- ifelse(test_set$height>cutoff_value, "Male", "Female") %>% factor(levels=levels(test_set$sex))
    list(method="Height cutoff",
         FPR = 1-specificity(y_hat, test_set$sex),
         TPR = sensitivity(y_hat, test_set$sex))
})
ggplot(data=height_cutoff, aes(x=FPR, y=TPR)) + geom_line() + geom_point() 
```
#Precision recall plot 
```{r echo=TRUE}
height_cutoff <- map_df(cutoffs, function (cutoff_value) {
    y_hat <- ifelse(test_set$height>cutoff_value, "Male", "Female") %>% factor(levels=levels(test_set$sex))
    list(method="Height cutoff",
         precision = precision(y_hat, test_set$sex),
         recall = sensitivity(y_hat, test_set$sex))
})
ggplot(data=height_cutoff, aes(x=recall, y=precision)) + geom_line() + geom_point() 
probs = seq(0.1, 0.9, by = 0.1)
guessing <- map_df(probs, function(p) {
  y_hat <- sample(c("Male", "Female"), length(test_index), replace=TRUE, prob=c(p, 1-p)) %>% factor(levels=levels(test_set$sex))
  list(method="Guessing",
       precision = precision(y_hat, test_set$sex),
       recall = sensitivity(y_hat, test_set$sex))
})
ggplot(data=height_cutoff, aes(x=recall, y=precision)) + geom_line() + geom_point() + geom_line(data=guessing, aes(x=recall, y=precision)) + geom_point(data=guessing, aes(x=recall, y=precision))
```

Excercises:
The following questions all ask you to work with the dataset described below.

The reported_heights and heights datasets were collected from three classes taught in the Departments of Computer Science and Biostatistics, as well as remotely through the Extension School. The Biostatistics class was taught in 2016 along with an online version offered by the Extension School. On 2016-01-25 at 8:15 AM, during one of the lectures, the instructors asked student to fill in the sex and height questionnaire that populated the reported_height dataset. The online students filled out the survey during the next few days, after the lecture was posted online. We can use this insight to define a variable which we will call type, to denote the type of student, inclass or online.

The code below sets up the dataset for you to analyze in the following exercises:

```{r}

library(dslabs)
library(dplyr)


data("reported_heights")

dat <- mutate(reported_heights, date_time = ymd_hms(time_stamp)) %>%
  filter(date_time >= make_date(2016, 01, 25) & date_time < make_date(2016, 02, 1)) %>%
  mutate(type = ifelse(day(date_time) == 25 & hour(date_time) == 8 & between(minute(date_time), 15, 30), "inclass","online")) %>%
  select(sex, type)

y <- factor(dat$sex, c("Female", "Male"))
x <- dat$type
dat %>% group_by(type) %>% summarize(prop_female = mean(sex == "Female"))

```
If you used the type variable to predict sex, what would the prediction accuracy be?
```{r}

  y_hat <- ifelse(dat$type == "inclass", "Male", "Female") 
  mean(y_hat == dat$sex)

```
Write a line of code using the table function to show the confusion matrix, assuming the prediction is y_hat and the truth is y.

```{r}
 y_hat <- ifelse(dat$type == "inclass", "Female", "Male") %>% factor(levels=levels(y))
 
 confusionMatrix(y_hat, y)
```

#Comprehension Check: Practice with Machine Learning
We will practice building a machine learning algorithm using a new dataset, iris, that provides multiple predictors for us to use to train. To start, we will remove the setosa species and we will focus on the versicolor and virginica iris species using the following code:

library(caret)
data(iris)
iris <- iris[-which(iris$Species=='setosa'),]
y <- iris$Species

The following questions all involve work with this dataset.
First let us create an even split of the data into train and test partitions using createDataPartition. The code with a missing line is given below:

```{r}
data(iris)
iris <- iris[-which(iris$Species=='setosa'),]
y <- iris$Species

set.seed(2)
test_index <- createDataPartition(y,times=1,p=0.5,list=FALSE)
# line of code
#Which code should be used in place of # line of code above?
#Answer:
 
test <- iris[test_index,]
train <- iris[-test_index,]
```


Answers are displayed within the problem Review
Q2
1/1 point (graded)
Next we will figure out the singular feature in the dataset that yields the greatest overall accuracy. You can use the code from the introduction and from Q1 to start your analysis.

Using only the train iris data set, which of the following is the singular feature for which a smart cutoff (simple search) yields the greatest overall accuracy?
```{r}
foo <- function(x){
	rangedValues <- seq(range(x)[1],range(x)[2],by=0.1)
	sapply(rangedValues,function(i){
		y_hat <- ifelse(x>i,'virginica','versicolor')
		mean(y_hat==train$Species)
	})
}
predictions <- apply(train[,-5],2,foo)
sapply(predictions,max)	
```

Using the smart cutoff value calculated on the training data, what is the overall accuracy in the test data?

```{r}
foo <- function(x){
	rangedValues <- seq(range(x)[1],range(x)[2],by=0.1)
	sapply(rangedValues,function(i){
		y_hat <- ifelse(x>i,'virginica','versicolor')
		mean(y_hat==test$Species)
	})
}
predictions <- apply(test[,-5],2,foo)
sapply(predictions,max)	
```
Notice that we had an overall accuracy greater than 96% in the training data, but the overall accuracy was lower in the test data. This can happen often if we overtrain. In fact, it could be the case that a single feature is not the best choice. For example, a combination of features might be optimal. Using a single feature and optimizing the cutoff as we did on our training data can lead to overfitting.

Given that we know the test data, we can treat it like we did our training data to see if the same feature with a different cutoff will optimize our predictions.

Which feature best optimizes our overall accuracy? - Petal.Width = 0.94

Now we will perform some exploratory data analysis on the data.sqrt(mean((y_hat-test_set$y)^2))

Notice that Petal.Length and Petal.Width in combination could potentially be more information than either feature alone.

Optimize the combination of the cutoffs for Petal.Length and Petal.Width in the train data and report the overall accuracy when applied to the test dataset. For simplicity, create a rule that if either the length OR the width is greater than the length cutoff or the width cutoff then virginica or versicolor is called. (Note, the F1 will be similarly high in this example.)

What is the overall accuracy for the test data now?

```{r}
petalLengthRange <- seq(range(train[,3])[1],range(train[,3])[2],by=0.1)
petalWidthRange <- seq(range(train[,4])[1],range(train[,4])[2],by=0.1)
cutoffs <- expand.grid(petalLengthRange,petalWidthRange)

id <- sapply(seq(nrow(cutoffs)),function(i){
	y_hat <- ifelse(train[,3]>cutoffs[i,1] | train[,4]>cutoffs[i,2],'virginica','versicolor')
	mean(y_hat==train$Species)
	}) %>% which.max

optimalCutoff <- cutoffs[id,] %>% as.numeric
y_hat <- ifelse(test[,3]>optimalCutoff[1] & test[,4]>optimalCutoff[2],'virginica','versicolor')
mean(y_hat==test$Species)


```

#Conditional probabilities. 

For classes k = 1...K
We have observed values x1,....,xp
For covariates X1,.....,Xp

This *does not imply* that the outcome y will take a specific value, but implies there is a specific probability.
We denote the conditional probabilities for different classes:
Pr(Y = k|X1=x1,...Xp=p) for k=1,....,K 

To avoid expressing X1,...,Xp and x1,...xp, we denote:
X=(X1,...,Xp)
x=(x1,...,xp)

So! The *conditional probability* of being in class K is denoted:
p<sub>k</sub>(x) = P(Y|X=x) for k=1,.....K

We will be using the notation p(x) to represent conditional probabilities as functions.

The goal here is to predict the highest probability of outcome belonging to a given class, among p1(x),p2(x),....p<sub>K</sub>(x)

E.g. How close the maximum probability is to 1:
max p<sub>k</sub>(x) is to 1 *or*
 
how close our estimate
$\widehat{p}$<sub>k</sub>(x) is to p<sub>k</sub>(x)

#Loss function
For fixed samples of data, specificity, sensitivity, accuracy and F1 are used to describe algorithm properties.
For continuous outcomes, though we define a loss function, which is denoted as $${(\widehat{Y} - Y)}^2$$
where $\widehat{Y}$ is predictor and Y actual outcome.

Since there are tests with multiple observation, the mean squared error is a mean to determine accuracy (an the actual accuracy for binary outcomes)
$$\frac{1}{N}\sum_{i=1}^{n}{(\widehat{Y}_i - Y_i)}^2$$
The general goal is to build an algorithm that minimises the loss and have a value close or equal to 0.
We want the algorithm that minimizes the average of the squared loss accross many random samples.
$$E\{\frac{1}{N}\sum_{i=1}^{n}{(\widehat{Y}_i - Y_i)}^2\}$$

e.g. the expectation of squared error - *expected loss*

##The main task of machine learning is to use data to estimate conditional probabilities 
$$f(x)≡E(Y|X=X)$$
for any set of features $${x=(x_1,...,x_p)}$$
#Excercisessqrt(mean((y_hat-test_set$y)^2))
We have a hypothetical population of 1 million individuals with the following conditional probabilities as described below:
The test is positive 85% of the time when tested on a patient with the disease (high sensitivity): 

P(test+|disease)=0.85
The test is negative 90% of the time when tested on a healthy patient (high specificity): 

P(test−|heathy)=0.90
The disease is prevalent in about 2% of the community: 

P(disease)=0.02
Here is some sample code to get you started:
set.seed(1)

```{r}
disease <- sample(c(0,1), size=1e6, replace=TRUE, prob=c(0.98,0.02))
test <- rep(NA, 1e6)
test[disease==0] <- sample(c(0,1), size=sum(disease==0), replace=TRUE, prob=c(0.90,0.10))
test[disease==1] <- sample(c(0,1), size=sum(disease==1), replace=TRUE, prob=c(0.15, 0.85))

print('#Q1: What is the probability that a test is positive?')
mean(test==1)
print('#Q2: What is the probability that an individual has the disease if the test is negative?')
mean(disease[test==0])
print('#Q3: What is the probability that you have the disease if the test is positive?')
mean(disease[test==1]==1)
print('#Q4: If the test is positive, what is the relative risk of having the disease?')
mean(disease[test==1]==1)/mean(disease==1)
```

We are now going to write code to compute conditional probabilities for being male in the heights dataset. Round the heights to the closest inch. Plot the estimated conditional probability 
P(x)=Pr(Male|height=x)
for each x. 
```{r}
library(dslabs)
data("heights")
hey <- heights %>% 
	mutate(height = round(height)) %>%
	group_by(height) %>%
	#summarize(p = mean(sex == "Male")) #%>%
	qplot(height, p, data =.)
```

In the plot we just made in Q1 we see high variability for low values of height. This is because we have few data points. This time use the quantile (\ 0.1,0.2,\dots,0.9 \)and the cut function to assure each group has the same number of points. Note that for any numeric vector x, you can create groups based on quantiles like this: cut(x, quantile(x, seq(0, 1, 0.1)), include.lowest = TRUE). 
```{r}
ps <- seq(0, 1, 0.1)
heights %>% mutate(g = cut(height, quantile(height, ps), include.lowest = TRUE))
  
```

You can generate data from a bivariate normal distrubution using the MASS package using the following code. 

```{r}
Sigma <- 9*matrix(c(1,0.5,0.5,1), 2, 2)
dat <- MASS::mvrnorm(n = 10000, c(69, 69), Sigma) %>%
	data.frame() %>% setNames(c("x", "y"))
plot(dat)
print("Using an approach similar to that used in the previous exercise, let's estimate the conditional expectations and make a plot. Part of the code has been provided for you: ")
ps <- seq(0, 1, 0.1)
dat %>% 
	mutate(g = cut(x, quantile(x, ps), include.lowest = TRUE)) %>%
  group_by(g) %>%
  summarize(y = mean(y), x = mean(x)) %>%
	qplot(x, y, data =.)
```

#Linear regression

```{r}
galton_heights <- GaltonFamilies %>% filter(childNum == 1 & gender == "male") %>% select (father, childHeight)
colnames(galton_heights)[colnames(galton_heights)=="childHeight"] <- "son"
y <- galton_heights$son
test_index <- createDataPartition(y, times = 1, p=0.5, list = FALSE)
train_set <- galton_heights %>% slice(-test_index)
test_set <- galton_heights %>% slice(test_index)
#Guessing
avg_height <- mean(train_set$son)
#Squared loss is: 
mean((avg_height - train_set$son)^2)
print("Getting intercept and slope:")
fit <-lm(son ~ father, data=train_set)
fit$coef
```

From the regression course:
$$f(x) = Pr(Y|X=x) = \beta_{0} + x*\beta_{1} => \widehat{f}(x) = 38 + 0.47x$$
Let's see how the sqared error results, instead of guessing:
```{r}
y_hat <- fit$coef[1] + fit$coef[2]*test_set$father
mean((y_hat - test_set$son)^2)
```

#The predict function
Replaces the formula for the regression line. 
```{r}
y_hat <- predict(fit, test_set)
```

#Excercises
Create a data set using the following code:
```{r}
set.seed(1)
n <- 100
Sigma <- 9*matrix(c(1.0, 0.5, 0.5, 1.0), 2, 2)
dat <- MASS::mvrnorm(n = 100, c(69, 69), Sigma) %>%
	data.frame() %>% setNames(c("x", "y"))
```

Use the caret package to partition the dataset into test and training sets of equal size. Train a linear model and calculate the RMSE. Repeat this exercise 100 times and report the mean and standard deviation of the RMSEs. (Hint: You can use the code shown in a previous course inside a call to replicate using a seed of 1.
```{r}

set.seed(1)
rmses <- replicate(100,  { 
  test_index <- createDataPartition(dat$y, times = 1, p=0.5, list = FALSE)
  train_set <- dat %>% slice(-test_index)
  test_set <- dat %>% slice(test_index)
  fit <-lm(y ~ x, data=train_set)
  y_hat <- predict(fit, test_set)
  rmse <- sqrt(mean((y_hat - test_set$y)^2)) 
  rmse
}) 
mean(rmses)
sd(rmses)
```

Now we will repeat the above but using larger datasets. Repeat the previous exercise but for datasets with n <- c(100, 500, 1000, 5000, 10000). Save the average and standard deviation of RMSE from the 100 repetitions using a seed of 1. Hint: use the sapply or map functions.

```{r}
set.seed(1)
map(c(100, 500, 1000, 5000, 10000), function(dataSetLength) {
  

  Sigma <- 9*matrix(c(1.0, 0.5, 0.5, 1.0), 2, 2)
  dat <- MASS::mvrnorm(n = dataSetLength, c(69, 69), Sigma) %>%
  data.frame() %>% setNames(c("x", "y"))
  rmses <- replicate(100,  { 
    test_index <- createDataPartition(dat$y, times = 1, p=0.5, list = FALSE)
    train_set <- dat %>% slice(-test_index)
    test_set <- dat %>% slice(test_index)
    fit <-lm(y ~ x, data=train_set)
    y_hat <- predict(fit, test_set)
    rmse <- sqrt(mean((y_hat - test_set$y)^2)) 
    rmse
  })
  
  print(paste("Items: ", toString(dataSetLength)))
  print(mean(rmses))
  print(sd(rmses))
})
```
Now repeat the exercise from Q1, this time making the correlation between x and y larger, as in the following code:

```{r}
set.seed(1)
n <- 100
Sigma <- 9*matrix(c(1.0, 0.95, 0.95, 1.0), 2, 2)
dat <- MASS::mvrnorm(n = 100, c(69, 69), Sigma) %>%
data.frame() %>% setNames(c("x", "y"))
set.seed(1)
rmses <- replicate(100,  { 
  test_index <- createDataPartition(dat$y, times = 1, p=0.5, list = FALSE)
  train_set <- dat %>% slice(-test_index)
  test_set <- dat %>% slice(test_index)
  fit <-lm(y ~ x, data=train_set)
  y_hat <- predict(fit, test_set)
  rmse <- sqrt(mean((y_hat - test_set$y)^2)) 
  rmse
}) 
mean(rmses)
sd(rmses)
```

Create a data set using the following code.
Note that y is correlated with both x_1 and x_2 but the two predictors are independent of each other, as seen by cor(dat).

Use the caret package to partition into a test and training set of equal size. Compare the RMSE when using just x_1, just x_2 and both x_1 and x_2. Train a linear model for each.

Which of the three models performs the best (has the lowest RMSE)?
```{r}
set.seed(1)
n <- 1000
Sigma <- matrix(c(1.0, 0.75, 0.75, 0.75, 1.0, 0.25, 0.75, 0.25, 1.0), 3, 3)
dat <- MASS::mvrnorm(n = 100, c(0, 0, 0), Sigma) %>%
	data.frame() %>% setNames(c("y", "x_1", "x_2"))

set.seed(1)
test_index <- createDataPartition(dat$y, times = 1, p = 0.5, list = FALSE)
train_set <- dat %>% slice(-test_index)
test_set <- dat %>% slice(test_index)

fit <- lm(y ~ x_1, data = train_set)
y_hat <- predict(fit, newdata = test_set)
sqrt(mean((y_hat-test_set$y)^2))

fit <- lm(y ~ x_2, data = train_set)
y_hat <- predict(fit, newdata = test_set)
sqrt(mean((y_hat-test_set$y)^2))

fit <- lm(y ~ x_1 + x_2, data = train_set)
y_hat <- predict(fit, newdata = test_set)
sqrt(mean((y_hat-test_set$y)^2))
```

Repeat the exercise from q6 but now create an example in which x_1 and x_2 are highly correlated.
```{r}
set.seed(1)
n <- 1000
Sigma <- matrix(c(1.0, 0.75, 0.75, 0.75, 1.0, 0.95, 0.75, 0.95, 1.0), 3, 3)
dat <- MASS::mvrnorm(n = 100, c(0, 0, 0), Sigma) %>%
	data.frame() %>% setNames(c("y", "x_1", "x_2"))

set.seed(1)
test_index <- createDataPartition(dat$y, times = 1, p = 0.5, list = FALSE)
train_set <- dat %>% slice(-test_index)
test_set <- dat %>% slice(test_index)

fit <- lm(y ~ x_1, data = train_set)
y_hat <- predict(fit, newdata = test_set)
sqrt(mean((y_hat-test_set$y)^2))

fit <- lm(y ~ x_2, data = train_set)
y_hat <- predict(fit, newdata = test_set)
sqrt(mean((y_hat-test_set$y)^2))

fit <- lm(y ~ x_1 + x_2, data = train_set)
y_hat <- predict(fit, newdata = test_set)
sqrt(mean((y_hat-test_set$y)^2))
```

#Logistic regression excercise
Define a dataset using the following code:

```{r}
set.seed(2)
make_data <- function(n = 1000, p = 0.5, 
				mu_0 = 0, mu_1 = 2, 
				sigma_0 = 1,  sigma_1 = 1){

y <- rbinom(n, 1, p)
f_0 <- rnorm(n, mu_0, sigma_0)
f_1 <- rnorm(n, mu_1, sigma_1)
x <- ifelse(y == 1, f_1, f_0)
  
test_index <- createDataPartition(y, times = 1, p = 0.5, list = FALSE)

list(train = data.frame(x = x, y = as.factor(y)) %>% slice(-test_index),
	test = data.frame(x = x, y = as.factor(y)) %>% slice(test_index))
}
dat <- make_data()
dat$train %>% ggplot(aes(x, color = y)) + geom_density()
set.seed(1)
delta <- seq(0, 3, len = 25)
res <- sapply(delta, function(d){
	dat <- make_data(mu_1 = d)
	fit_glm <- dat$train %>% glm(y ~ x, family = "binomial", data = .)
	y_hat_glm <- ifelse(predict(fit_glm, dat$test) > 0.5, 1, 0) %>% factor(levels = c(0, 1))
	mean(y_hat_glm == dat$test$y)
})
qplot(delta, res)
```
#Smoothing
##Problem with one predictor
Let's estimate the time trend in the popular vote from the 2008 election.
```{r}
data("polls_2008")
qplot(day, margin, data=polls_2008)
```
###Let's predict y given the day x.
This is the mathematical model for the observed poll margin Y.
$Y_i=f(x_i)+\epsilon_i$
The conditional expectation is:
f(x)=E(Y|X=x)
Linear regression shows:
```{r}
day_margin_fit <-lm(margin ~ day, data=polls_2008)
polls_2008_scatter_data <- polls_2008 %>% mutate(resid=resid(day_margin_fit), color=ifelse(resid(day_margin_fit) >0,"+","-"), fitted=fitted(day_margin_fit))
ggplot(aes(day, margin, color=color), data=polls_2008_scatter_data) + geom_point() + geom_abline(intercept = day_margin_fit$coefficients[1], slope=day_margin_fit$coefficients[2])
    
#span <- 7 fit <- with(polls_2008, ksmooth(day, margin, x.points = day, kernel="normal", bandwidth = span))

#polls_2008 %>% mutate(smooth = fit$y) %>% ggplot(aes(day, margin)) + geom_point(size = 3, alpha = .5, color = "grey") + geom_line(aes(day, smooth), color="red")
```
